The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-11-27 manuscript/manuscript.qmd
Conditionally hide introduction heading
See https://github.com/quarto-dev/quarto-cli/issues/7593
Rearrange sections and clean up outline

---
title: "The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society"
short-title: "China's ONGO Law and International Civil Society"
date: "2023-09-25"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@student.gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
---

@@

# Introduction

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. The law requires all international NGOs (INGOs) to either register representative offices (ROs) or file temporary activities to lawfully operate in mainland China.As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness INGOs.

TODO: Andrew - update and coordinate the introductory paragraphs with the theoretical framing below

*Meng's notes/suggestions*: to highlight the theoretical significance in the very beginning (just listing the idea pool for moment, not concise enough):

 1. A frequently approach to close the space of INGOs is passing restrictive laws/institutional approach. [@DeMattee:2019]. To further understanding this global phenomena , China is an important case to explore the impacts such restrictive laws have on the operation of INGOs. Such legal restrictions can be either direct (denial of legal access/cannot register) or covert (red-tape, requiring the approval of PSUs, or the uncertainty of having to filing TAs case by case)

 2. *de facto* vs. *de jure* (incorporating the paragraph below) and differentiated treatment for more vs. less contentious INGOs. Authoritarian -> arbitrary and varied application of the law -> different *operating space*

 3. Laws and policies as a *policy tool* and behavioral shifit: to tame INGO to be more cooperative to government. The requirement of PSU and Chinese partner to re-direct INGOs to be more authority friendly.

The impacts on international civil society  we examine include not only access or representation of INGOs in China, but also whether there is change of behaviors [@WuNie:2021]. Mention the concept of *flexibility (in the sense of adjusting to change)* too?

@@

# Closing civic space

## The global constriction of civic space

## The "closing space" phenomenon

INGO regulations are implemented and enforced differentially - difference in de jure and de facto. This differential enforcement is one mechanism to allow host governments to harness the aspects of INGO programming that align with their domestic policy preferences while limiting INGO behavior that could threaten regime stability.

## Closing civic space in China

@@

3. Laws and policies as a *policy tool* and behavioral shifit: to tame INGO to be more cooperative to government. The requirement of PSU and Chinese partner to re-direct INGOs to be more authority friendly.

@@

# Legal background

@@

# Theoretical background and expectations

@@

This paper aims to fill these gaps by empirically examining how the "closing space" hypothesis applies to China in the years after the ONGO Law took effect. China provides an excellent setting for testing these theories because China's promulgation of restrictive INGO law as an authoritarian jurisdiction fits the observed global trends in recent years, thus can serve as an representative case to deepen scholarly understanding around these themes. Also, the level of ambiguity of the legal framework gives ample room for arbitrary discretion and conflicting attitudes by authorities[@Jia:2017], enabling us to analyze rules in form (*de jure*) as well as rules in use (*de facto*).

## The "closing space" phenomenon

@@

## De jure vs. de facto implementation of civil society regulations

@@

## China's INGO as a policy tool

## Empirical expectations

@@

- Stats + case study: **Flexibility**
- Stats + case study: **Contentiousness**
- Case study: **Relationship with PSU and government** (stats/descriptive bc lots of chambers of commerce

@@

@@

::: callout-warning
**Still TODO!**
:::

>>>>

---
title: "The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society"
short-title: "China's ONGO Law and International Civil Society"
date: "2023-11-27"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@student.gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
---

@@

::: {.content-visible when-format="html"}
# Introduction {.hide}
:::

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. The law requires all international NGOs (INGOs) to either register representative offices (ROs) or file temporary activities to lawfully operate in mainland China. As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness INGOs.

We make three general theoretical contributions in this paper:

1. Expansion of work using institutional approaches to examine the regulation and restriction of civil society - [@DeMattee:2019]. To further understanding this global phenomena , China is an important case to explore the impacts such restrictive laws have on the operation of INGOs. Such legal restrictions can be either direct (denial of legal access/cannot register) or covert (red-tape, requiring the approval of PSUs, or the uncertainty of having to filing TAs case by case)

2. Looking at how civil society laws act as a policy tool that regimes can leverage to shift NGO behavior: to tame INGO to be more cooperative to government. The requirement of PSU and Chinese partner to re-direct INGOs to be more authority friendly - The impacts on international civil society we examine include not only access or representation of INGOs in China, but also whether there is change of behaviors [@WuNie:2021]. Mention the concept of *flexibility (in the sense of adjusting to change)* too?

3. Exploring differentiated application of these laws - *de facto* vs. *de jure* and differentiated treatment for more vs. less contentious INGOs. Authoritarian -> arbitrary and varied application of the law -> different *operating space*

TODO: New cool thing here is that we can look at de facto implementation in greater detail. Existing studies use things like V-Dem's core civil society index or civil society repression index to capture the general regulatory environment, but changes in those scores aren't necessarily tied to specific laws. Here we can see that differential enforcement/de facto/de jure stuff directly

@@

# Closing civic space in China - legal background thing

@@

@@

---

@@

# Theoretical background

@@

This paper aims to fill these gaps by empirically examining how the "closing space" hypothesis applies to China in the years after the ONGO Law took effect. China provides an excellent setting for testing these theories because China's promulgation of restrictive INGO law as an authoritarian jurisdiction fits the observed global trends in recent years, thus can serve as an representative case to deepen scholarly understanding around these themes. Also, the level of ambiguity of the legal framework gives ample room for arbitrary discretion and conflicting attitudes by authorities [@Jia:2017], enabling us to analyze rules in form (*de jure*) as well as rules in use (*de facto*).

## The "closing space" phenomenon - Institutional approach to closing space

@@

## Regulations as policy tool to shape behavior

## De jure vs. de facto implementation of civil society regulations - Differential enforcement

@ChaudhryHeiss:2022 and the divergence in de jure and de facto regulations---the implementation doesn't have to match the written law, and often the written law can serve as a "weapon hanging on the wall that never fires" that serves as a warning to potentially uncooperative NGOs [@Kozenko:2015]. See also [@ToeplerZimmerFrohlich:2020; @Plantan:2022]
INGO regulations are implemented and enforced differentially - difference in de jure and de facto. This differential enforcement is one mechanism to allow host governments to harness the aspects of INGO programming that align with their domestic policy preferences while limiting INGO behavior that could threaten regime stability.

@@

# Empirical expectations

@@

- Stats + case study: **Flexibility**
- Stats + case study: **Contentiousness**
- Case study: **Relationship with PSU and government** (stats/descriptive bc lots of chambers of commerce)

@@

Paragraph justifying why the case study is needed

Education, Environment, lower expected operating space (flexibility?)

So Issue areas can predict some level of flexibility/restrictions INGOs receive, but that is not the full story, the relations with the government (one more dimension/variable) also plays important roles, see TNC vs. Greenpeace

@@
